---
title: "Lab 4"
author: "Erica Bishop"
date: "2023-05-09"
output: html_document
---

```{r packages, include = FALSE}
library(tidytext)
library(tidyverse)
library(tidymodels)
library(textrecipes)
library(naivebayes)
library(discrim) # naive-bayes
library(kernlab) #kernlab engine
```

Lab 4 Assignment: Due May 9 at 11:59pm

```{r load_data, message=FALSE}

urlfile ="https://raw.githubusercontent.com/MaRo406/EDS-231-text-sentiment/main/data/climbing_reports_model_dat.csv"
incidents_df<-readr::read_csv(url(urlfile))

```

### 1. Select another classification algorithm

I'll use a support vector machine classification model becuase SVM's are efficient where there are a lot of featuers (i.e. a lot of tokens - as in this text analysis). 

### 2. Conduct an initial out-of-the-box model fit on the training data and prediction on the test test data.  Assess the performance of this initial model.

```{r split data}
#set seed for reproducibility
set.seed(567)

#transform data so category is two classes 
incidents2class <- incidents_df %>% 
  mutate(fatal = factor(if_else(is.na(Deadly),
                                "non-fatal",
                                "fatal")))

#split data to train model
incidents_split <- initial_split(incidents2class)

incidents_train <- training(incidents_split)
incidents_test <- testing(incidents_split)

```

```{r set model specifications}

#set recipe
recipe <- recipe(fatal ~ Text,
       data = incidents_train) %>% 
  step_tokenize(Text) %>% #turning each word in text into a different feature
  step_tokenfilter(Text, max_tokens = 1000) %>%  #max_tokens argument limits to just most frequent words (number of features you want)
  step_tfidf(Text) #term frequency individual document frequency takes into account tfdf metric


#use default engine and tuning parameters
svm_spec <- svm_linear() %>% 
  set_engine("kernlab") %>% 
  set_mode("classification")

#bundle into a workflow
svm_wf <- workflow() %>% 
  add_recipe(recipe) %>% 
  add_model(svm_spec)

```

```{r fit message=FALSE} 

svm_fit <- svm_wf %>% 
  fit(data = incidents_train)
  
```

3. Select the relevant hyperparameters for your algorithm and tune your model.

4. Conduct a model fit using your newly tuned model specification.  What are the terms most highly associated with non-fatal reports?  What about fatal reports? 

5. Predict fatality of the reports in the test set.  Compare this prediction performance to that of the Naive Bayes and Lasso models.